Easy as Pi, Volume II: On the Goodput and Energy of Network Coding Schemes in the Raspberry Pi

### 2016-08-14 conclusions.tex
more changes

\label{sec:conclusions}
Given the usefulness of the \ac{Raspi} as a low-complex
processing node in large-scale networks and network coding
techniques against state-of-the-art routing, we provide
a performance evaluation of network coding schemes focusing on
processing speed and energy consumption for two \ac{Raspi} models.
The evaluation includes algorithms that exploit both \ac{SIMD}
instructions and multicore capabilities of the \ac{Raspi} 2. Our
measurements show that processing speeds of more than 80 Mbps and 800
Mbps are attainable for the \ac{Raspi} model 1 and 2, respectively,
for a wide range of network coding configurations and maintaining
a processing energy below 1 nJ/bit (or even an order of magnitude lower)
in similar configurations. For the use of multithreading, we quantify
processing gains ranging from 2$\times$ for $g = 16$ to 13$\times$ for
$g = 1024$ when employing 4 threads each in a different core. Future
work in the use of \ac{Raspi} devices will focus on considering: (i)
the performance of the \ac{Raspi} in scenarios with synthetic packet
losses, (ii) wireless networks where real packet losses can occur
and (iii) other topologies such as broadcast or
cooperative scenario to compare with theoretical results in order to
evaluate the performance of different network codeswith the \ac{Raspi}.

>>>>

\label{sec:conclusions}
Given the usefulness of the \ac{Raspi} as a low-complex
processing node in large-scale networks and network coding
techniques against state-of-the-art routing, we provide
a performance evaluation of network coding schemes focusing on
processing speed and energy consumption for two \ac{Raspi} models.
The evaluation includes algorithms that exploit both \ac{SIMD}
instructions and multicore capabilities of the \ac{Raspi} 2. Our
measurements show that processing speeds of more than 80 Mbps and 800
Mbps are attainable for the \ac{Raspi} model 1 and 2, respectively,
for a wide range of network coding configurations and maintaining
a processing energy below 1 nJ/bit (or even an order of magnitude lower)
in similar configurations. For the use of multithreading, we quantify
processing gains ranging from 2$\times$ for $g = 16$ to 13$\times$ for
$g = 1024$ when employing 4 threads each in a different core. Future
work in the use of \ac{Raspi} devices will focus on considering: (i)
the performance of the \ac{Raspi} in scenarios with synthetic packet
losses, (ii) wireless networks where real packet losses can occur
and (iii) other topologies such as broadcast or
cooperative scenario to compare with theoretical results in order to
evaluate the performance of different network codes with the \ac{Raspi}.

### 2016-08-14 measurements.tex
some corrections
more changes

With the testbed, methodology and setups from the previous sections, we
proceed to obtain the measurements for the \ac{Raspi} 1 and 2 platforms.
We consider the following set of parameters for our study:
For all the codes, we use $g = [16,\ 32,\ 64,\ 128,\ 256,\ 512]$
and $q = [2,\ 2^8]$. For the non-optimized multicore implementations
and cases when the generation size is varied $B = 1600$ bytes.
We consider a setup where we only vary the packet size,
$B = [64,\ 128,\ 256,\ 512,\ 1024,\ 2048,\ 4096,\ 8192,\ 16384,\ 32768
,\ 65536,\ 131072]$ bytes with a generation size fixed on $g = [16,\ 128]$
to see computing load that the \ac{Raspi} present in the low and high
packet size regimes. We considered $B = 1536$ bytes for the optimized
multicore implementation. For \ac{SRLNC}, we considered
$d = [0.02,\ 0.1]$. For \ac{TSNC}, we considered excess packets,
$e = [8,\ 16]$, so that the budget is simply $b = g + e$. In all our
measurement reports, to simplify their review, we first presents the
results for the \ac{Raspi} 1 and later continue with the \ac{Raspi} 2.

@@

\hline
\ac{Raspi} Model & $I_{idle,avg}$ (A)& $I_{proc,avg}$ (A) & $P_{avg}$ (W) \\
\hline
\hline
    \ac{Raspi} 1 &      0.320        &            0.360    & 0.200\\
\hline
    \ac{Raspi} 2 &      0.216        &            0.285    & 0.345\\
\hline
\end{tabular}
\vspace{0.2cm}
\label{tab:average_power_table}
\end{table}
%
From Table~\ref{tab:average_power_table}, we notice that the power expenditure
for both models is almost the same. Thus, the energy behaviour is mostly
dependent on the goodput trends since the power is just an scaling constant.
%
\subsection{Energy per bit}
With power and goodput measurements, we compute the energy per bit of
the previously mentioned cases as described in \eqref{eq:energy_per_bit}.
The trends of the energy per bit are simply the inverse of the goodput since
both metrics are related by an inverse law. As made with the goodput, we
separate the results descriptions according to the operation carried by the
\ac{Raspi}: Encoding and decoding. Besides this, we differentiate between
the models in our study. The energy per bit consumption removes the dependency
on the amount of packets helping to normalize the results and indicating
energy consumption on a fair basis for all the configurations and coding
operations involved in the study.

@@

To review the performance of \ac{NC} in a multicore architecture using the
\ac{Raspi}, the authors implemented the algorithm described in
Section~\ref{sub:implementation-multicore} on the \ac{Raspi} 2 model B, which
features four ARMv7 cores in a Broadcom BCM2836 \ac{SOC} with a 900 MHz clock.
Each core has a 32 KiB L1 data cache and 32 KiB L1 instruction cache. The cores
share a 512 KiB L2 cache. All the measured results, including the baseline
results, were obtained with NEON-enabled code adapted from the Fifi library
\cite{fifi}. The \ac{Raspi} 2 has a NEON extension instruction set which
provides an 128-bit \ac{SIMD} instructions that speed the computations.
Fig.~\ref{enc_dec1024}, \ref{enc_dec128} and \ref{enc_dec16} show the
encoding and decoding goodput in MB per second for different generation sizes,
$g = [1024, 128, 16]$ respectively. For $g = [128, 16]$, the displayed results
are the mean values over 1000 measurements, while for $g=1024$, they are mean
values over 100 measurements. The size of each coded packet was fixed to
$1536$~bytes since that is the typical size of an Ethernet frame. The blocked
operations were performed dividing the matrices in squared sub-blocks of $16,\
32\ ,\ 64,\ldots,\ 1024$ operands (words in the Galois field) in height and
width. The figures show only block sizes of $16 \times 16$, and $128 \times 128$
operands since with bigger block sizes, the operands do not fit in cache.
Several testcases are considered and detailed.

>>>>

%With the testbed, methodology and setups from the previous sections, we
With the methodology and setups from the previous sections, we
proceed to obtain the measurements for the \ac{Raspi} 1 and 2 devices.
We consider the following set of parameters for our study:
For all the codes, we use $g = [16,\ 32,\ 64,\ 128,\ 256,\ 512]$
and $q = [2,\ 2^8]$. For the non-optimized multicore implementations
and cases when the generation size is varied $B = 1600$ bytes.
We consider a setup where we only vary the packet size,
$B = [64,\ 128,\ 256,\ 512,\ 1024,\ 2048,\ 4096,\ 8192,\ 16384,\ 32768
,\ 65536,\ 131072]$ bytes with a generation size fixed on $g = [16,\ 128]$
to see computing load that the \ac{Raspi} present in the low and high
packet size regimes. We considered $B = 1536$ bytes for the optimized
multicore implementation. For \ac{SRLNC}, we considered
$d = [0.02,\ 0.1]$. For \ac{TSNC}, we considered excess packets,
$e = [8,\ 16]$, so that the budget is simply $b = g + e$. In all our
measurement reports, to simplify their review, we first presents the
results for the \ac{Raspi} 1 and later continue with the \ac{Raspi} 2.

@@

\hline
\ac{Raspi} Model & $I_{idle,avg}$ (A)& $I_{proc,avg}$ (A) & $P_{avg}$ (W) \\
\hline
\hline
    \ac{Raspi} 1 &      0.320        &            0.360    & 0.200\\
\hline
    \ac{Raspi} 2 &      0.216        &            0.285    & 0.345\\
\hline
\end{tabular}
\vspace{0.2cm}
\label{tab:average_power_table}
\end{table}
%
From Table~\ref{tab:average_power_table}, we notice that the power expenditure
for both models is almost the same. Thus, the energy behaviour is mostly
dependent on the goodput trends since the power is just a scaling constant.
%
\subsection{Energy per bit}
With power and goodput measurements, we compute the energy per bit of
the previously mentioned cases as described in \eqref{eq:energy_per_bit}.
The trends of the energy per bit are simply the inverse of the goodput since
both metrics are related by an inverse law. As made with the goodput, we
separate the results descriptions according to the operation carried by the
\ac{Raspi}: Encoding and decoding. Besides this, we differentiate between
the models in our study. The energy per bit consumption removes the dependency
on the amount of packets helping to normalize the results and indicating
energy consumption on a fair basis for all the configurations and coding
operations involved in the study.

@@

To review the performance of \ac{NC} in a multicore architecture using the
\ac{Raspi}, we implemented the algorithm described in
Section~\ref{sub:implementation-multicore} on the \ac{Raspi} 2 model B, which
features four ARMv7 cores in a Broadcom BCM2836 \ac{SOC} with a 900 MHz clock.
Each core has a 32 KiB L1 data cache and 32 KiB L1 instruction cache. The cores
share a 512 KiB L2 cache. All the measured results, including the baseline
results, were obtained with
NEON enabled.
%NEON-enabled code adapted from the Fifi library \cite{fifi}.
The \ac{Raspi} 2 has a NEON extension instruction set which
provides an 128-bit \ac{SIMD} instructions that speed the computations.
Fig.~\ref{enc_dec1024}, \ref{enc_dec128} and \ref{enc_dec16} shows the
encoding and decoding goodput in MB per second for different generation sizes,
$g = [1024, 128, 16]$ respectively. For $g = [128, 16]$, the displayed results
are the mean values over 1000 measurements, while for $g=1024$, they are mean
values over 100 measurements. The size of each coded packet was fixed to
$1536$~bytes since that is the typical size of an Ethernet frame. The blocked
operations were performed dividing the matrices in squared sub-blocks of $16,\
32\ ,\ 64,\ldots,\ 1024$ operands (words in the Galois field) in height and
width. The figures show only block sizes of $16 \times 16$, and $128 \times 128$
operands since with bigger block sizes, the operands do not fit in cache.
Several testcases are considered and detailed.
